Characterising Optimisations to Memory Access Patterns using Architecture Independent Program Features

### 2020-01-17 paper/paper.tex
add GEM analysis, working on abstract

	This paper addresses the challenge of profiling a code base to help programmers effectively optimise their codes for high performance.
	Current methods of analysing program behaviour have largely relied on performance counters on various architectures to provide architecture dependent metrics of code behaviour. 
	However, if workload characteristics are tied to architectural features specific to a particular system, these analysis frameworks may not generalise to alternative or future systems.
	Considering the trend towards heterogeneous computing, we work towards a framework to guide developers in optimising codes for any given hardware target using the architecture independent workload characterisation (AIWC) tool -- specifically in profiling memory access patterns of code. 
	Metrics must be easily interpreted and conceptually linked to patterns in hardware-specific memory access optimisations.
	To this end we propose, implement and analyse the addition of two new metrics to AIWC and present a new approach to profile spatial locality intrinsic to parallel codes.
	The improvements suggested to AIWC will help HPC developers better understand memory access patterns of complex codes, a crucial step in guiding optimisation of codes for arbitrary hardware targets.
	\end{abstract}
	\maketitle

@@

	% As all kernels access memory on word (32-bit) boundaries, there is no effect in dropping upto two bits in the PSL metric...
	\begin{figure*}[h!]
		\centering
		\begin{tikzpicture}
		\begin{axis}[symbolic x coords={0,1,2,3,4,5,6,7,8,9,10},
		width=0.7\textwidth,
		height=0.5\textwidth,
		xtick={0,1,2,3,4,5,6,7,8,9,10},
		xticklabel style={text height=2ex},
		ymin=0,
		ylabel={$n$-bits dropped Parallel Spatial Locality}, xlabel={\# of bits skipped},
		ylabel style={at={(axis description cs:0.05,.5)},anchor=south},
		legend entries={\texttt{GEM}, \texttt{NW*}, \texttt{SRAD Kernel 1},\texttt{SRAD Kernel 2},
			\texttt{CSR}, \texttt{BFS*}, \texttt{LUD Diagonal}, \texttt{LUD Internal}, \texttt{LUD Perimeter}},
		legend style={at={(0.55,0.75)},anchor=west}
		]

@@

	% TODO

	\subsection{Dense Linear Algebra: Lower-upper decomposition (LUD)}

>>>>

	This paper addresses the challenge of profiling a code base to help programmers effectively optimise their codes for high performance.
	The Architecture Independent Workload Characterisation Tool (AIWC) gathers features of OpenCL programs used to understand and predict program performance.

	Current methods of analysing program behaviour have largely relied on performance counters on various architectures to provide architecture dependent metrics of code behaviour. 
	However, if workload characteristics are tied to architectural features specific to a particular system, these analysis frameworks may not generalise to alternative or future systems.
	Considering the trend towards heterogeneous computing, we work towards a framework to guide developers in optimising codes for any given hardware target using the architecture independent workload characterisation (AIWC) tool -- specifically in profiling memory access patterns of code.

	Metrics must be easily interpreted and conceptually linked to patterns in hardware-specific memory access optimisations.
	To this end we propose, implement and analyse the addition of two new metrics to AIWC and present a new approach to profile spatial locality intrinsic to parallel codes.
	The improvements suggested to AIWC will help HPC developers better understand memory access patterns of complex codes, a crucial step in guiding optimisation of codes for arbitrary hardware targets.
	\end{abstract}
	\maketitle

@@

	% As all kernels access memory on word (32-bit) boundaries, there is no effect in dropping upto two bits in the PSL metric...
	\begin{figure*}[h!]
		\centering
		\begin{tikzpicture}
		\begin{axis}[symbolic x coords={0,1,2,3,4,5,6,7,8,9,10},
		width=0.7\textwidth,
		height=0.5\textwidth,
		xtick={0,1,2,3,4,5,6,7,8,9,10},
		xticklabel style={text height=2ex},
		ymin=0,
		ylabel={$n$-bits dropped Parallel Spatial Locality}, xlabel={\# of bits skipped},
		ylabel style={at={(axis description cs:0.05,.5)},anchor=south},
		legend entries={\texttt{GEM}, \texttt{NW*}, \texttt{SRAD Kernel 1},\texttt{SRAD Kernel 2},
			\texttt{CSR}, \texttt{BFS*}, \texttt{LUD Diagonal}, \texttt{LUD Internal}, \texttt{LUD Perimeter}, \texttt{TDM}},
		legend style={at={(0.55,0.75)},anchor=west}
		]

@@

	% TODO
	The OpenDwarfs benchmark GEM computes the electrostatic potential of a biomolecule by calculating the sum of charges contributed by all
	This is an embarrassingly parallel problem. Each OpenCL work-item operates on a single surface vertex, finding the electrostatic potential generated by looping through every atom in the biomolecule in a global order. 
	The computation pattern is highly regular and memory accesses are perfectly synchronised. Atom data is accessed consecutively, with all work-items simultaneously accessing each single atom's data.
	The parallel spatial locality metric reflects this pattern of efficient memory utilisation (single loads from global memory servicing all OpenCL work-items). The recorded parallel spatial locality approaches the theoretical limit of 0 (0.0124 at 10 bits dropped). This indicates that almost all memory accesses made by the OpenCL kernel are perfectly synchronised between OpenCL threads.
	%TODO: Clarify above sentence: I mean efficient memory fetches (one load from global memory services all threads) -- not reduced usage of memory.
	\subsection{Dense Linear Algebra: Lower-upper decomposition (LUD)}
